An Empirical Analysis of the Role of Energy in Economic Growth

### 2013-05-24 EconGrowth.Rnw
Fixed the f enduse-to-intake information. Changed to f_intake/wholesale.

\noindent where $C_{fuel}$ is the consumption rate of the fuel (in units of short tons/year for coal, barrels/year for oil and petroleum, ft$^3$/year for dry natural gas and natural gas plant liquids, and m$^3$/year for biomass/fuel wood) and $h_{{fuel}}$ is the heat content of the fuel (in kJ/short ton for coal, kJ/barrel for oil, kJ/ft$^3$ for dry natural gas and natural gas plant liquids, and kJ/kg for biomass/fuel wood). Fuel wood density (166.67~kg/m$^3$ **** Caleb: verify this number and source ****) was calculated from **** and ****, both obtained from the \citet{FAO:2013c}.

@@

\noindent where the value of $Q_{m,workers}$ for the U.S. ($\sim$ 4200 kcal/day-worker) was taken from a time series of kcal/day-person for workers for 1980--2011 \citep{Warr:2012a}. $Q_{m,workers}$ for other countries was taken from the \citet[``\texttt{Dietary Energy Cons.~Countries}'' tab]{FAO:2013b}. The value for $N_{workers}$ was obtained from the labor time series for each economy as discussed in Section \ref{sec:Labor_Data}.

@@

\noindent $\eta_{food \rightarrow work, animals}$ and $\eta_{food \rightarrow work, workers}$ are animal and worker efficiency of converting food into energy, respectively. The animal feed-to-work conversion efficiency ($\eta_{feed \rightarrow work, animals}$ = 0.4) was taken from \citet{Warr:2012a} and assumed equal across all economies. The food-to-work efficiency for workers ($\eta_{food \rightarrow work, workers}$ = 0.2) was taken from \citet[Table~B.1]{Ayres:2010ug}. Following \citet[p.~315]{Ayres:2010ug}, an additional factor ($f_{enduse/intake}$) was included to account for workers' food energy that sustains the body's normal function as opposed to providing energy for physical labor \citep{Wirsenius:2000vi} **** Caleb needs to find a detailed source (table, equation, number, page number) in Wirsenius for this approach ****. $f_{enduse/intake}$ varies by country (Table \ref{tab:f_enduse_intake}). A similar enduse-to-intake ratio could also be applied to animal work, but it was unavailable from \citet{Wirsenius:2000vi}.\footnote{The error in exergy caused by neglecting $f_{enduse/intake}$ for animals is very small, because animal muscle exergy is a small fraction of the total energy supply to any economy, with a maximum of 0.41\% for China in 1992.}

\begin{table} \caption{Food enduse-to-intake ratios ($f_{enduse/intake}$) for all countries.} \label{tab:f_enduse_intake}
  \begin{center}
    \begin{tabular}{r c c c c c c c c c}
      Country & US & UK & JP & CN & ZA & SA & IR & TZ & ZM \\
      \hline
      $f_{enduse/intake}$ & 0.62 & 0.65 & 0.76 & 0.76 & 0.87 & 0.72 & 0.72 & 0.72 & 0.72\\
    \end{tabular}
  \end{center}
\end{table}

>>>>

\noindent where $C_{fuel}$ is the consumption rate of the fuel (in units of short tons/year for coal, barrels/year for oil and petroleum, ft$^3$/year for dry natural gas and natural gas plant liquids, and m$^3$/year for biomass/fuel wood) and $h_{{fuel}}$ is the heat content of the fuel (in kJ/short ton for coal, kJ/barrel for oil, kJ/ft$^3$ for dry natural gas and natural gas plant liquids, and kJ/kg for biomass/fuel wood). Fuel wood density of 300~kg/m$^3$ was taken from \citet{Ragland:2013wp} at the low end of the range of density of dry wood and bark.

@@

\noindent where the value of $Q_{m,workers}$ is the ``the amount of food available for human consumption'' (i.e., at the wholesale level) taken from the \citet[``\texttt{Dietary~Energy~Cons.~Countries}'' tab]{FAO:2013b}. The value for $N_{workers}$ was obtained from the labor time series for each economy as discussed in Section~\ref{sec:Labor_Data}.

@@

\noindent $\eta_{food \rightarrow work, animals}$ and $\eta_{food \rightarrow work, workers}$ are animal and worker efficiency of converting food into energy, respectively. The animal feed-to-work conversion efficiency ($\eta_{feed \rightarrow work, animals}$ = 0.4) was taken from \citet{Warr:2012a} and assumed equal across all economies. The food-to-work efficiency for workers ($\eta_{food \rightarrow work, workers}$ = 0.2) was taken from \citet[Table~B.1]{Ayres:2010ug}. Following \citet[p.~315]{Ayres:2010ug} and methodology from \citet{Warr:2012a}, an additional factor ($f_{intake/wholesale}$) was included to account for ``wastage and losses of food in the household, e.g. during storage, in preparation and cooking, as plate-waste, or quantities fed to domestic animals and pets'' \citep{FAO:2013b}. Values of $f_{intake/wholesale}$ were calculated from \citet[Table~3.3 for wholesale and Table~3.4 for intake]{Wirsenius:2000vi}. $f_{intake/wholesale}$ varies by country (Table \ref{tab:f_intake_wholesale}). A similar intake-to-wholesale ratio could also be applied to animal work, but data to calculate the ratio was unavailable from \citet{Wirsenius:2000vi}.\footnote{The error in exergy caused by neglecting $f_{intake/wholesale}$ for animals is very small, because animal muscle exergy is a small fraction of the total energy supply to any economy, with a maximum of 0.41\% for China in 1992. It could be argued that $f_{intake/wholesale}$ for animals is likely to be closer to 1.0 than $f_{intake/wholesale}$ for humans.}

\begin{table} \caption{Food intake-to-wholesale ratios ($f_{intake/wholesale}$) for all countries.} \label{tab:f_intake_wholesale}
  \begin{center}
    \begin{tabular}{r c c c c c c c c c}
      Country & US & UK & JP & CN & ZA & SA & IR & TZ & ZM \\
      \hline
      $f_{enduse/intake}$ & 0.62 & 0.65 & 0.76 & 0.76 & 0.87 & 0.72 & 0.72 & 0.72 & 0.72\\
    \end{tabular}
  \end{center}
\end{table}
